Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-05-04 ForegroundSimulations/reply_to_referee_report.txt

Referee Report

Reviewer's Comments:

While 21cm cosmology promises a tremendous amount of new information in both astrophysics and cosmology, the weakness of the signal combined with the dominance of foregrounds poses daunting requirements on instrument characterization. The work presented in this manuscript directly addresses some of the most urgent yet poorly understood aspects of 21 cm instruments, so it is highly valuable to the field and timely for the deployment and future expansion of the HERA project. Therefore, I recommend this work for publication, with minor modifications to address my comments below:

Major comments:

1) It is not clear if the design of HERA dish is good enough or not. In sections 2 and 6, the authors convey the message that HERA dish design is good enough (even "optimal"), whereas in section 5 it is clear that more dish measurements and possible design tweaks are needed. It is important to present a more coherent picture on this front.

Response: The intention of this paper and companion papers is to illustrate that HERA is basing its design after comprehensive analysis of its effects on detectability of the EoR signal. Though many parameters of the HERA design are close to being final, efforts are ongoing to fine-tune the instrument based on measurements. I have rephrased statements to emphasize this picture.

2) Section 4.1 needs to be more explicit regarding the assumptions and limitations in the HFSS simulation, especially those related to effects investigated in this work. For example, does the HFSS consider multiple reflections between the feed and the dish? How about between dishes?

Response: Added a sentence at the end of first paragraph under section 4.1 that the models include mutiple reflections between the dish and the feed but not those between dishes.

3) The choice of arbitrary pointing in section 5.1 undermines the credibility of this work. I understand that the choice of pointing may not be very important here, but an arbitrary choice with no further explanation is not acceptable. Fig. 6 seems to suggest certain sky directions (around RA hour 17) can have very different impacts on the results.

Response: I have now noted in section 5.1 that choice of pointing is immaterial for HERA as long as pointings that have the Galactic plane overhead (in RA range 16-19 hours) are avoided. In fact, no one specific pointing was chosen only to demonstrate that HERA is robust to all pointings except for a narrow range in RA.

4) For figures 2 and 5, it seems that presenting curves for all 8 baseline lengths conveys little more information than presenting just the shortest and longest baselines. Maybe less can be more in this case?

Response: I have noted in section 5.1 that foreground contamination depends on baseline length, orientation, pointing direction and spatial structure of emitting source. Hence, I chose a range of baseline lengths and orientations. I used all unique baseline lengths to demonstrate that all baselines individually will be sensitive to the EoR signal.

Minor comments:

Abstract:
1) "with high significance": this sounds vague. Is there a sigma value or a signal-to-noise ratio (SNR) to quote?

Response: Unless thermal noise is included it is difficult to quote a SNR. Here I have assumed data is not thermal noise limited but only limited by foregrounds and systematics studied in this paper. In such a scenario the significance of detection again varies anywhere from 10 to more than 1000.

2) "All baselines...are capable": does this mean each baseline individually can make the detection, or all of them combined? Since this work focuses on beam chromaticity and just assumes sufficient SNR, it is not clear if such statement regarding detection is truly justifiable in the scope of this work.

Response: It is meant to imply all baselines individually can make the detection as long as they are not thermal-noise limited but only limited by foreground and instrumental systematics studied here. Hence the sentence is now prefixed with "Under these conditions" which refers to data that is limited by foregrounds and such systematics discussed in this paper.

Section 1:
1) "there are inherent differences in spatial isotropy and spectral smoothness between the EoR signal and the foregrounds": explicitly describe the difference in one sentence?

Response: Added a sentence as suggested.

2) "whereas the EoR signal has spherical symmetry due to its isotropy": two issues here. First, the 'whereas' is a strange transition. The wedge bounded by horizon limits versus the isotropy of EoR signal do not prove any point as far as I can tell. The wedge is there even if the foreground were also isotropic, and EoR signal is subject to the horizon limits just like the foreground. Second, EoR signal is statistically isotropic, so saying the signal is spherically symmetric might be misleading for readers not familiar with the subject.

Response: Reworded to indicate foregrounds are inside the wedge due to the horizon limits and smooth spectrum while the EoR power spectrum is spherically symmetric due to statistical isotropy and fluctuating in its spectrum.

Section 2:

1) "currently under construction on the South African SKA site": nit-picking here (everywhere?), but it might sound confusing why HERA is built on the SKA site. Maybe just omit the SKA site information since it is not relevant?

Response: Replaced SKA site with "South African Karoo Radio Astronomy Reserve".

2) "strike an optimal balance between sensitivity and systematics": this is directly related to my major comments #1. First of all, even if that were the case, it is not quite appropriate to state it as a matter of fact in the introduction, since it is what this work is setting out to prove. Secondly, it seems clear in this work that the dish design has not truly achieved the optimum. Sections 5.2.1 and 5.2.2 explicitly state that 'measurements are under way" and "the HERA collaboration constantly pursues improvement of its dish design".

Response: Removed the word "optimal". The balance between sensitivity and systematics was initially studied in DeBoer and Parsons (2015) - a HERA memo - to which a citation has now been added. This study was specific to choice of dish size and preliminary estimates of specifications for dish based on sensitivity and systematics. Our current work along with companion papers are a much more comprehensive effort to further reinforce findings of that preliminary study. It is in this context that sentence was used. It is true that the dish design is not finalized yet and is still being fine-tuned for optimum performance. Hence, removal of the word "optimum" is consistent with other statements about measurements and fine-tuning of the dish design.

3) I think it is never explained in this section that HERA 19 is already built, and full HERA with 320 elements is being proposed?

Response: The last paragraph in section 2 does mention this.

Section 4:
1) It would be helpful to explain in half a sentence what kind of software HFSS is. Is it a commercial simulation software, or a customized HERA software?

Response: Added a brief description

2) The authors go into great detail as to what the materials are used in simulation, which leaves the reader wondering: are they the same as what are actually used in the dish? For example, "the backplane is modeled as an aluminum surface", but is it actually an aluminum surface? A nice picture of an actual HERA dish may help here.

Response: TBD

3) Can the authors briefly comment on the relative importance of point sources versus the diffuse emission in the simulation? Are the point sources relevant for the 14 m baselines? What about the 60 m ones?

Response: TBD

Section 5:
1) "The equation for delay spectrum": explicit \eqref (that is, identify the equation)

Response: Explicitly refer to equations 1 and 2 as suggested

2) "by convolving a Blackman-Harris window with itself": I might be wrong here, but Fig. 1 looks like a multiplication in frequency and convolution in delay domain? If it is squared in the delay domain, where do the ripples come from in the BH*BH curve in the delay domain since the BH curve is smooth?

Response: It is definitely a multiplication in delay domain as equation 10 indicates. This is what leads to the squaring and increase in dynamic range in delay domain. However, it is not a squaring of the same exact BH delay domain curve (gray) shown in figure 1 (bottom). This is because after convolving in frequency the resultant window must fit in a 100 MHz band end-to-end. An auto-convolution of the BH curve shown in figure 1 (which already already occupies 100 MHz end-to-end) will result in a curve that takes up 200 MHz end-to-end. Hence a new BH curve with appropriate width is chosen such that after auto-convolution, the resultant curve occupies 100 MHz. In other words, the black curve in top panel is not obtained by convolving gray curve with itself but from a slightly different BH curve of different (smaller) width with itself. This is the reason the null-like dips in delay domain are not aligned between the two. I have added a few sentences to describe this subtlety.

The ripples seen could be because of the aforementioned adjustments in frequency domain. Also being at the level of 10**-8 to 10**-5 they are insignificant compared to the overall dynamic range. In other words, even if the gray curve had ripples at this level, in a log scale plot they will not appear as ripples when superimposed on the tail which is at the level of 10 to 10**4. Thus it will not be surprising that the gray curve will still appear smooth despite these ripples.

3) If the square of the Blackman-Harris window works so much better, why not the third or fourth power?

Response: With higher order powers of the delay domain response, they start getting broader in delay and narrower in frequency. This results in loss of thermal noise senstivity due to loss of effective bandwidth. And viewing from a Fourier domain perspective, due to loss of delay resolution, sensitivity to large scale structure will reduce. In conclusion, it is possible to use higher order powers but depending on the application, this may result in loss of sensitivity especially on large scales. In this paper, we found that squaring was sufficient to achieve the dynamic range required for our purposes. I have added a small note highlighting this trade-off.

4) Fig 1: the axis labels for the top plot have non math mode fonts, which are inconsistent with all other plots.

Response: Corrected as suggested

5) Fig 1: 'as a result' is a strange transition since the previous sentence is talking about epsilon.

Response: Removed the phrase "As a result"

6) "cause further contamination": it is not clear what 'further' is being compared to. What is the original contamination?

Response: Added a sentence to indicate this contamination is in addition to that due to other spectral structures in the foregrounds and the instrument.

7) "small amplitude differences": Fig. 5 (a) seems to show as much as 2 orders of magnitude difference between the two EoR models?

Response: Qualified as relatively small differences when only the relevant modes $k_\parallel > k_p$ are considered.

Section 6:
1) The transition between sections 5 and 6 can use a few more sentences. It is not very clear how the dynamic range relates to the dish's reflectometry.

Response: Added a transition sentence at the beginning of the first paragraph in section 6.

2) Fig. 6: comment on the dark stripe around RA hour 17?

Response: The last paragraph in section 6 discusses this. I have also added a sentence describing it in the caption of figure 6.

Section 7:
1) "robustly eliminate these systematics further": robustly and further sound strange together.

Response: Changed to "effectively reduce these systematics further"

### 2016-05-04 ForegroundSimulations/ms.tex
Make most of the changes suggested by referee
From 9e3f5a4 (master):
=====================

In ms.tex,

(*) Make most of the changes suggested by referee

In eor.bib,

(*) Add citation to HERA memo

Unaccounted for systematics from foregrounds and instruments can severely limit the sensitivity of current experiments from detecting redshifted 21~cm signals from the Epoch of Reionization (EoR). Upcoming experiments are faced with a challenge to deliver more collecting area per antenna element without degrading the data with systematics. This paper and its companions show that dishes are viable for achieving this balance using the Hydrogen Epoch of Reionization Array (HERA) as an example. Here, we specifically identify spectral systematics associated with the antenna power pattern as a significant detriment to all EoR experiments which causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated design specifications on these reflections to prevent further EoR signal degradation. We show HERA will not be impeded by such spectral systematics and demonstrate that even in a conservative scenario that does not perform removal of foregrounds, HERA will detect EoR signal in line-of-sight $k$-modes, $k_\parallel \gtrsim 0.2\,h$~Mpc$^{-1}$, with high significance. All baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

The most significant challenge to low frequency EoR observations arises from the extremely bright Galactic and extragalactic foreground synchrotron emission which are $\sim 10^4$ times stronger than the desired EoR signal \citep{dim02,ali08,ber09,ber10,gho12}. However, there are inherent differences in spatial isotropy and spectral smoothness between the EoR signal and the foregrounds \citep[see, e.g.,][]{fur04b,mor04,zal04,san05,fur06,mcq06,mor06,wan06,gle08}.

When expressed in the coordinate system of power spectrum measurements described by the three-dimensional wavenumber ($k$), the foreground emission is restricted to a wedge-shaped region commonly referred to as the {\it foreground wedge} \citep{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,dil14} bounded by the horizon limits \citep{par12b}, whereas the EoR signal has spherical symmetry due to its isotropy. The region of $k$-space excluding the {\it foreground wedge} is commonly referred to as the {\it EoR window} since it is expected to be relatively free of foreground contamination.

@@

Along with companion papers \citep{ewa16,neb16,pat16}, we focus on determining whether larger antenna elements can be viable for EoR experiments. In this paper, we explore the impact of chromaticity in the antenna power pattern on extending foreground power beyond the {\it wedge}. While highlighting the importance of such instrumental systematics on future EoR experiments, we demonstrate that the current HERA instrument design will exceed these specifications to detect the EoR signal with high significance. This will hold true when data is limited by foregrounds and systematics even in the most conservative scenario which involves neither foreground subtraction nor optimal analysis techniques but only relies on a simple spectral weighting scheme.

@@

\citet{deb16} describe the HERA instrument in detail while \citet{pob14} explored the range of astrophysical parameters that HERA can probe. We provide a summary of the instrument currently under construction on the South African SKA site.

@@

HERA is deploying 14~m fixed zenith-pointing parabolic dishes that strike an optimal balance between sensitivity and systematics \citep[][and this paper]{ewa16,neb16,pat16}. The large collecting area of a HERA element yields $\approx 5$ times the sensitivity of an MWA tile and more than 20 times that of a PAPER element.

@@

In this study, we use the HERA-19 hexagonal array layout, a small subset of the planned layout that are currently in use. Without loss of generality, the analysis and results in this paper will apply to the full proposed layout as well.

@@

The High Frequency Structural Simulator (HFSS) was used to model the dish and its angular response used in this study. The HFSS model used prime focus optics with a 14~m faceted parabola with a spar f/D ratio of 0.32.  The model has a 1~m central hole in the aluminum surface which is filled with a dielectric material similar to dry soil. The feed used a full PAPER dipole inside of a cylindrical backplane, the backplane is modeled as an aluminum surface. For the metal parts of the dipole, the discs were modeled as aluminum at the actual size, and the arms and terminals were modeled as copper. Dielectric stand-offs and supporting members were included. For the calculations, one pair of arms was excited using a modal port. These models cover a frequency range of 90--210~MHz in intervals of 1~MHz \citep{deb16}.

@@

Our all-sky foreground model is the same as the one in \citet{thy15a}. It consists of diffuse emission \citep{deo08} and point sources. The latter is obtained from a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz with a mean spectral index of -0.83. The diffuse sky model has an angular resolution of 13\farcm 74 and a spectral index estimated for every pixel.

@@

The equation for delay spectrum defines the interplay between foregrounds and antenna power pattern, and the mapping between geometric phases of baselines and delays. We discuss the spillover of foreground power beyond the horizon delay limits caused by the chromatic nature of the antenna power pattern.

@@

Fig.~\ref{fig:window-functions} shows the {\it Blackman-Harris} window (gray) and  the window in equation~\ref{eqn:bhw2} (black) in frequency (top) and their power responses in delay (bottom). In its Fourier domain, this window has a response obtained by squaring the {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB with $\epsilon\approx 42\%$ and $w_\textrm{ne} \approx 2.88$.

\begin{figure}[htb]
  \centering
  % \includegraphics[width=\linewidth]{window_function_modifications.eps}
  \includegraphics[width=\linewidth]{f1.eps}
  \caption{Choices for spectral weighting functions, $W(f)$ ({\it top}) and their delay power spectrum responses, $|\widetilde{W}(\tau)|^2$ ({\it bottom}). The gray curves correspond to a {\it Blackman-Harris} window while those in black correspond to a {\it Blackman-Harris} window convolved with itself. The overall sensitivity of the former is $\epsilon \approx 50\%$ of a rectangular window while that of the latter is $\epsilon \approx 42\%$. As a result, the main lobe of the convolved window function response is slightly wider than that of a {\it Blackman-Harris} window. However, the sidelobes from the convolved window function are suppressed by more than ten orders of magnitude relative to that of a nominal {\it Blackman-Harris} window.}
  \label{fig:window-functions}
\end{figure}

@@

In HERA-19 array layout, there are 30 unique baseline vectors and 8 unique baseline lengths. Fig.~\ref{fig:asm-dps-beam-chromaticity-baselines} shows the delay power spectra of foregrounds on the 8 unique baseline lengths obtained with the aforementioned models for power pattern at arbitrary pointings on the sky. In all these panels, the full-band foreground delay power spectra obtained with achromatic, {\it Airy} and chromatic simulated beam patterns are shown in black, red, and blue respectively. The brightening of foreground power near the horizon limits (vertical dotted lines) due to the {\it pitchfork} effect \citep{thy15a,thy15b} is prominently seen in all cases. A clear broadening of spillover-wings outside the horizon limits is seen with increasing chromaticity as the beam is changed from the achromatic to the {\it Airy} to the chromatic simulated model. For instance, the spillover from foreground delay power spectrum obtained with chromatic beam pattern is restricted to $|k_\parallel| \lesssim 0.2\,h$~Mpc$^{-1}$, with the {\it Airy} pattern it is restricted to $|k_\parallel| \lesssim 0.15\,h$~Mpc$^{-1}$, while with the achromatic beam it is restricted to $|k_\parallel| \lesssim 0.12\,h$~Mpc$^{-1}$ even on longest baseline lengths. It is also noted that despite the extreme dynamic range of the spectral weighting function employed, the tail of the foreground spillover at $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ with an {\it Airy} beam pattern is many orders of magnitude higher than that using an achromatic beam while that from the simulated chromatic beam is even higher than from an {\it Airy} pattern by a few orders of magnitude. Thus, the foreground spillover shown is truly limited by intrinsic spectral features in the antenna beam patterns.

\begin{figure*}[htb]
  \centering
  % \includegraphics[width=\linewidth]{asm_foreground_eor_beam_chromaticity_fullband_bhw2.0.eps}
  \includegraphics[width=\linewidth]{f2.eps}
  \caption{Foreground delay power spectra in units of K$^2 (h^{-1}$~Mpc$)^3$ on unique baseline lengths of HERA-19 at arbitrary sky pointings. The baseline length and orientation (anti-clockwise from East) are annotated at the top right corner of each panel. Black, red and blue curves correspond to delay power spectra obtained with achromatic, {\it Airy} and chromatic simulated antenna beams respectively. The achromatic beam has no spectral structure, the simulated chromatic beam has maximum chromaticity while a nominal {\it Airy} pattern has intermediate level of chromaticity. The {\it pitchfork} effect is clearly visible as peaks around the horizon limits. With increase in chromaticity of the antenna beam the foreground spillover beyond the {\it foreground wedge} becomes progressively worse - the extent of foreground-spillover wings beyond the horizon limits (vertical dotted lines) and the amplitude of spillover beyond $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ is most severe for the chromatic simulated beam, intermediate for an {\it Airy} pattern and negligible for an achromatic beam.}
  \label{fig:asm-dps-beam-chromaticity-baselines}
\end{figure*}

@@

Reflections shift the measured foreground power to higher modes in $\tau$ (or in $k_\parallel$) and thus cause further contamination in these critical higher $k_\parallel$-modes. These delay shifts introduce ripples in the spectrum. The net chromaticity in the measurements is the product of spectral structure arising out of the intrinsic nature of foreground emission, $I(\hat{\boldsymbol{s}},f)$, the baseline- and position-dependent frequency structure of the geometric phases, $e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}$, the spectral features in the antenna power pattern, $A(\hat{\boldsymbol{s}},f)$, besides any other spectral structures in the instrument unaccounted for. In $P(\boldsymbol{b},k_\parallel)$ defined in the Fourier domain, these factors have a convolving effect.

@@

Both EoR models were employed in this analysis to reduce the dependence of design specifications on any single model. However, we found similar results except for minor differences caused by small amplitude differences between the two EoR models. Hence, we only present results from using EoR model 1 at 150~MHz.

@@

We estimate the EoR-foreground dynamic range (ratio of EoR signal and foreground powers) for both EoR models in the presence of antenna beam chromaticity, using HERA as an example. In order to avoid signal evolution across the entire band, we use subbands in which EoR signal evolution is not expected to be significant. These subbands with $B_\textrm{eff}=10$~MHz are defined by spectral weights in equation~\ref{eqn:bhw2}. These subbands are centered at 150~MHz and 170~MHz to estimate EoR-foreground dynamic range at $z \approx 8.47$ and $z \approx 7.36$ respectively.

@@

However, in all four cases considered here, except when the Galactic center appears overhead in the RA range roughly in the range 16-19 hr, the worst case signal-foreground ratio lies above unity in most of this parameter space. It demonstrates that HERA should detect the EoR signal with a very high signal-foreground ratio on all baselines over a majority of the pointings.

@@

By accounting for these instrumental systematics, if data is not limited by thermal noise, we demonstrate that HERA should detect EoR in $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ with high significance with a simple spectral weighting even under a conservative analysis criterion that does not involve {\it foreground removal} strategies. All baselines of HERA-19 will not only detect but also help in filtering highly likely EoR models.

We are also investigating advanced strategies for foreground removal that will robustly eliminate these systematics further. Advances in calibration, foreground modeling and subtraction will improve the performance listed here. More optimal analysis techniques and further design improvements will make HERA not only clear the obstacles limiting first-generation experiments but also become transformational for next generation of low-frequency cosmology experiments.

>>>>

Unaccounted for systematics from foregrounds and instruments can severely limit the sensitivity of current experiments from detecting redshifted 21~cm signals from the Epoch of Reionization (EoR). Upcoming experiments are faced with a challenge to deliver more collecting area per antenna element without degrading the data with systematics. This paper and its companions show that dishes are viable for achieving this balance using the Hydrogen Epoch of Reionization Array (HERA) as an example. Here, we specifically identify spectral systematics associated with the antenna power pattern as a significant detriment to all EoR experiments which causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated design specifications on these reflections to prevent further EoR signal degradation. We show HERA will not be impeded by such spectral systematics and demonstrate that even in a conservative scenario that does not perform removal of foregrounds, HERA will detect EoR signal in line-of-sight $k$-modes, $k_\parallel \gtrsim 0.2\,h$~Mpc$^{-1}$, with high significance. {\bf Under these conditions,} all baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

The most significant challenge to low frequency EoR observations arises from the extremely bright Galactic and extragalactic foreground synchrotron emission which are $\sim 10^4$ times stronger than the desired EoR signal \citep{dim02,ali08,ber09,ber10,gho12}. However, there are inherent differences in spatial isotropy and spectral smoothness between the EoR signal and the foregrounds \citep[see, e.g.,][]{fur04b,mor04,zal04,san05,fur06,mcq06,mor06,wan06,gle08}. {\bf For instance, the foregrounds have smooth spectrum and are not isotropic, whereas the EoR signal manifests with a fluctuating spectrum and statistical isotropy.}

When expressed in the coordinate system of power spectrum measurements described by the three-dimensional wavenumber ($k$), the foreground emission is restricted to a wedge-shaped region commonly referred to as the {\it foreground wedge} \citep{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,dil14} bounded by the horizon limits \citep{par12b}, {\bf owing to its smooth spectrum}. {\bf On the other hand, the EoR power spectrum has spherical symmetry due to statistical isotropy and non-smooth spectrum}. The region of $k$-space excluding the {\it foreground wedge} is commonly referred to as the {\it EoR window} since it is expected to be relatively free of foreground contamination.

@@

Along with companion papers \citep{ewa16,neb16,pat16}, we focus on determining whether larger antenna elements can be viable for EoR experiments. In this paper, we explore the impact of chromaticity in the antenna power pattern on extending foreground power beyond the {\it wedge}. While highlighting the importance of such instrumental systematics on future EoR experiments, we demonstrate that the HERA instrument design {\bf planned} will detect the EoR signal with high significance. This will hold true when data is limited by foregrounds and systematics even in the most conservative scenario which involves neither foreground subtraction nor optimal analysis techniques but only relies on a simple spectral weighting scheme.

@@

\citet{deb16} describe the HERA instrument in detail while \citet{pob14} explored the range of astrophysical parameters that HERA can probe. We provide a summary of the instrument currently under construction {\bf in the South African Karoo Radio Astronomy Reserve}.

@@

HERA is deploying 14~m fixed zenith-pointing parabolic dishes that {\bf aim to} strike a balance between sensitivity and systematics \citep[][and this paper]{deb15,ewa16,neb16,pat16}. The large collecting area of a HERA element {\bf will yield} $\approx 5$ times the sensitivity of an MWA tile and more than 20 times that of a PAPER element.

@@

In this study, we use the HERA-19 hexagonal array layout, a small subset of the planned layout that {\bf is} currently in use. Without loss of generality, the analysis and results in this paper will apply to the full proposed layout as well.

@@

The High Frequency Structural Simulator (HFSS), {\bf a commercial finite element method solver for electromagnetic structures from Ansys,} was used to model the dish and its angular response used in this study. The HFSS model used prime focus optics with a 14~m faceted parabola with a spar f/D ratio of 0.32.  The model has a 1~m central hole in the aluminum surface which is filled with a dielectric material similar to dry soil. The feed used a full PAPER dipole inside of a cylindrical backplane, the backplane is modeled as an aluminum surface. For the metal parts of the dipole, the discs were modeled as aluminum at the actual size, and the arms and terminals were modeled as copper. Dielectric stand-offs and supporting members were included. For the calculations, one pair of arms was excited using a modal port. These models cover a frequency range of 90--210~MHz in intervals of 1~MHz {\bf and include multiple reflections between the dish and the feed but not those between dishes} \citep{deb16}.

@@

{\bf We use the foreground model of \citet{thy15a} wherein diffuse and point sources were shown to affect foreground contamination differently. The model for diffuse emission is obtained from \citet{deo08} and that for point sources} is obtained from a combination of the NRAO VLA Sky Survey \citep[NVSS;][]{con98} at 1.4~GHz and the Sydney University Molonglo Sky Survey \citep[SUMSS;][]{boc99,mau03} at 843~MHz with a mean spectral index of -0.83. The diffuse sky model has an angular resolution of 13\farcm 74 and a spectral index estimated for every pixel.

@@

{\bf Equations~\ref{eqn:obsvis} and \ref{eqn:delay-transform} define} the interplay between foregrounds and antenna power pattern, and the mapping between geometric phases of baselines and delays. We discuss the spillover of foreground power beyond the horizon delay limits caused by the chromatic nature of the antenna power pattern.

@@

Fig.~\ref{fig:window-functions} shows the {\it Blackman-Harris} window (gray) and  the window in equation~\ref{eqn:bhw2} (black) in frequency (top) and their power responses in delay (bottom). In its Fourier domain, this window has a response obtained by squaring {\bf a} {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB with $\epsilon\approx 42\%$ and $w_\textrm{ne} \approx 2.88$.

\begin{figure}[htb]
  \centering
  % \includegraphics[width=\linewidth]{window_function_modifications.eps}
  \includegraphics[width=\linewidth]{f1.eps}
  \caption{Choices for spectral weighting functions, $W(f)$ ({\it top}) and their delay power spectrum responses, $|\widetilde{W}(\tau)|^2$ ({\it bottom}). The gray curves correspond to a {\it Blackman-Harris} window while those in black correspond to a {\it Blackman-Harris} window convolved with itself. The overall sensitivity of the former is $\epsilon \approx 50\%$ of a rectangular window while that of the latter is $\epsilon \approx 42\%$. The main lobe of the convolved window function response is slightly wider than that of a {\it Blackman-Harris} window. However, the sidelobes from the convolved window function are suppressed by more than ten orders of magnitude relative to that of a nominal {\it Blackman-Harris} window.}
  \label{fig:window-functions}
\end{figure}

@@

{\bf It must be noted that in order to obtain this modified window function over a band of 100~MHz after convolution, the original {\it Blackman-Harris} window used in the convolution must occupy a band narrower than 100~MHz. This is why the modified window has a wider response in delay domain. This loss of delay-domain resolution, or equivalently, narrower effective bandwidth results in a slight reduction of sensitivity towards noise as well as small $k$-modes. While higher order powers could be used instead of the squared response we employ, there is a trade-off between the dynamic range and sensitivity. A squared response was found to be sufficient for our purposes.}

In HERA-19 array layout, there are 30 unique baseline vectors and 8 unique baseline lengths. {\bf \citet{thy15a} showed that foreground contamination depends not only on the baseline length but also on its orientation, pointing direction on the sky, and whether the source of emission is diffuse or compact}. Fig.~\ref{fig:asm-dps-beam-chromaticity-baselines} shows the delay power spectra of foregrounds on the 8 unique baseline lengths obtained with the aforementioned models for power pattern at {\bf certain chosen} pointings on the sky. {\bf It will be shown later in \S\ref{sec:baseline-RA} that choice of pointing is immaterial for HERA as long as pointings with the Galactic plane overhead are avoided}. In all these panels, the full-band foreground delay power spectra obtained with achromatic, {\it Airy} and chromatic simulated beam patterns are shown in black, red, and blue respectively. The brightening of foreground power near the horizon limits (vertical dotted lines) due to the {\it pitchfork} effect \citep{thy15a,thy15b} is prominently seen in all cases. A clear broadening of spillover-wings outside the horizon limits is seen with increasing chromaticity as the beam is changed from the achromatic to the {\it Airy} to the chromatic simulated model. For instance, the spillover from foreground delay power spectrum obtained with chromatic beam pattern is restricted to $|k_\parallel| \lesssim 0.2\,h$~Mpc$^{-1}$, with the {\it Airy} pattern it is restricted to $|k_\parallel| \lesssim 0.15\,h$~Mpc$^{-1}$, while with the achromatic beam it is restricted to $|k_\parallel| \lesssim 0.12\,h$~Mpc$^{-1}$ even on longest baseline lengths. It is also noted that despite the extreme dynamic range of the spectral weighting function employed, the tail of the foreground spillover at $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ with an {\it Airy} beam pattern is many orders of magnitude higher than that using an achromatic beam while that from the simulated chromatic beam is even higher than from an {\it Airy} pattern by a few orders of magnitude. Thus, the foreground spillover shown is truly limited by intrinsic spectral features in the antenna beam patterns.

\begin{figure*}[htb]
  \centering
  % \includegraphics[width=\linewidth]{asm_foreground_eor_beam_chromaticity_fullband_bhw2.0.eps}
  \includegraphics[width=\linewidth]{f2.eps}
  \caption{Foreground delay power spectra in units of K$^2 (h^{-1}$~Mpc$)^3$ on unique baseline lengths of HERA-19 at {\bf certain chosen} sky pointings. The baseline length and orientation (anti-clockwise from East) are annotated at the top right corner of each panel. Black, red and blue curves correspond to delay power spectra obtained with achromatic, {\it Airy} and chromatic simulated antenna beams respectively. The achromatic beam has no spectral structure, the simulated chromatic beam has maximum chromaticity while a nominal {\it Airy} pattern has intermediate level of chromaticity. The {\it pitchfork} effect is clearly visible as peaks around the horizon limits. With increase in chromaticity of the antenna beam the foreground spillover beyond the {\it foreground wedge} becomes progressively worse - the extent of foreground-spillover wings beyond the horizon limits (vertical dotted lines) and the amplitude of spillover beyond $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ is most severe for the chromatic simulated beam, intermediate for an {\it Airy} pattern and negligible for an achromatic beam.}
  \label{fig:asm-dps-beam-chromaticity-baselines}
\end{figure*}

@@

Reflections shift the measured foreground power to higher modes in $\tau$ (or in $k_\parallel$) and thus cause contamination in these critical $k_\parallel$-modes {\bf which is in addition to that already present due to spectral structures in the foregrounds and the instrument}. These delay shifts introduce ripples in the spectrum. The net chromaticity in the measurements is the product of spectral structure arising out of the intrinsic nature of foreground emission, $I(\hat{\boldsymbol{s}},f)$, the baseline- and position-dependent frequency structure of the geometric phases, $e^{-i2\pi f\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}}$, the spectral features in the antenna power pattern, $A(\hat{\boldsymbol{s}},f)$, besides any other spectral structures in the instrument unaccounted for. In $P(\boldsymbol{b},k_\parallel)$ defined in the Fourier domain, these factors have a convolving effect.

@@

Both EoR models were employed in this analysis to reduce the dependence of design specifications on any single model. However, we found similar results except for minor differences caused by {\bf relatively} small amplitude differences between the two EoR models {\bf for $|k_\parallel|>k_\textrm{p}$}. Hence, we only present results from using EoR model 1 at 150~MHz.

@@

{\bf The HERA antenna beam model used accounts for chromaticity introduced due to reflections between the dish and the feed but not between dishes. Assuming the latter is insignificant,} we estimate the EoR-foreground dynamic range (ratio of EoR signal and foreground powers) {\bf achievable with HERA} for both EoR models in the presence of antenna beam chromaticity. In order to avoid signal evolution across the entire band, we use subbands in which EoR signal evolution is not expected to be significant. These subbands with $B_\textrm{eff}=10$~MHz are defined by spectral weights in equation~\ref{eqn:bhw2}. These subbands are centered at 150~MHz and 170~MHz to estimate EoR-foreground dynamic range at $z \approx 8.47$ and $z \approx 7.36$ respectively.

@@

However, in all four cases considered here, except when the Galactic {\bf plane} appears overhead {\bf roughly between 16 and 19 hours of RA}, the worst case signal-foreground ratio lies above unity in most of this parameter space. It demonstrates that HERA should detect the EoR signal with a very high signal-foreground ratio on all baselines over a majority of the pointings.

@@

By accounting for these instrumental systematics, if data is not limited by thermal noise, we demonstrate that HERA should detect EoR in $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ with high significance with a simple spectral weighting even under a conservative analysis criterion that does not involve {\it foreground removal} strategies. All baselines of HERA-19 will not only {\bf individually} detect but also help in filtering highly likely EoR models.

We are also investigating advanced strategies for foreground removal that will {\bf effectively reduce} these systematics further. Advances in calibration, foreground modeling and subtraction will improve the performance listed here. More optimal analysis techniques and further design improvements will make HERA not only clear the obstacles limiting first-generation experiments but also become transformational for next generation of low-frequency cosmology experiments.
